\documentclass[a4paper, twoside]{report}

%% Language and font encodings
\usepackage[english]{babel}
\usepackage[utf8]{inputenc}
\usepackage[T1]{fontenc}

%% Sets page size and margins
\usepackage[a4paper,top=3cm,bottom=2cm,left=3cm,right=3cm,marginparwidth=2cm]{geometry}

%% Useful packages
\usepackage{amsmath}
\usepackage{xspace}
\usepackage{graphicx}
\usepackage{array, booktabs}
\usepackage{listings}
\usepackage[dvipsnames]{xcolor}
\usepackage{textcomp}
\usepackage{wrapfig}
\usepackage{float}
\usepackage[chapter]{minted}
\usepackage{caption}
\usepackage{subcaption}
\usepackage[colorinlistoftodos]{todonotes}
\usepackage[colorlinks=true, allcolors=blue]{hyperref}

% Separation logic style sheet
\usepackage{sl}
\usepackage{gillian}

\setlength{\parindent}{0em}
\setlength{\parskip}{1em}
\providecommand*{\listingautorefname}{Listing}

\title{Gillian Debugging: Stepping Up, Over and Into User Experience}
\author{Nat Karmios}

\begin{document}

% Text in \autoref references
\def\chapterautorefname{\normalcolor{Chapter}\color{blue}}
\def\sectionautorefname{\normalcolor{\S}\kern-0.7ex\color{blue}}
\def\subsectionautorefname{\normalcolor{\S}\kern-0.7ex\color{blue}}
\def\figureautorefname{\normalcolor{Figure}\color{blue}}
\def\lstnumberautorefname{\normalcolor{line}\color{blue}}
\newcommand{\sacha}[1]{\todo[color=green]{S: {#1}}}

\input{title/title.tex}

\begin{abstract}
This project improves upon previous work to provide complex debugging functionality to
Gillian, a multi-language real-world program analysis platform built by the
Verified Software group at Imperial College London, which is parametric on the memory
model of the target language.
The previous solution was restricted to linear, step-by-step execution by its
adherence to the Debug Adapter Protocol. This project, by reconfiguring Gillian's log
trace structure, implementing branched execution, extending the DAP, and
implementing a custom UI via webview, has resulted in the creation of a rich debugging interface for debugging Gillian program verification.
This interface, created for Visual Studio Code, enables customised debugging of symbolic execution traces, providing, for example, the ability to isolate and switch between various execution branches, as well as to step through and inspect Gillian's unification process, which is essential for debugging specifications in verification. This has greatly improved the overall usability of Gillian, bringing it closer to real-world users, such as Imperial College students and general developers.
\end{abstract}

\renewcommand{\abstractname}{Acknowledgements}
\begin{abstract}
My thanks go to Sacha Ayoun for being a keen eye and a guiding hand throughout
the course of the project.

I'd also like to thank Professor Philippa Gardner and Petar Maksimović,
whose teaching and project proposals formed the academic path I walk, to this
project and beyond.

Finally, I give my love to Douglas, who has been an eternally patient rock at my
side.
\end{abstract}

% \listoftodos{}
\tableofcontents

\input{content/intro.tex}
\input{content/background.tex}
\input{content/current.tex}
\input{content/improvements.tex}
\input{content/evaluation.tex}
\input{content/conclusion.tex}
\input{content/ethics.tex}
\input{content/refs.tex}
\input{content/appendix.tex}

\end{document}
