%!TEX root = ../../main.tex
%chktex-file 18

\section{Rethinking Gillian's Internal Mechanisms}

\subsection{Making the Interpreter Branch-Aware}%
\label{sec:interpreter-branching}

A particularly interesting consequence of implementing the given logging
structure was the inclusion of \texttt{branch\_case} fields, both in
\texttt{"cmd"} and \texttt{"cmd\_result"} reports (see \autoref{lst:branch-case}
for the relevant type). As can be seen in \autoref{fig:log-structure-desired},
the multiple results of a command that branches (such as an if/else), and the
relevant commands that directly follow, are associated with each other via their
\texttt{branch\_case} fields.

\texttt{branch\_case}s could then be used as a mechanism for dictating branched
execution. Each state configuration type has been extended with a field of type
\texttt{branch\_path} (equivalent to a \texttt{branch\_case list}),  denoting
the `path' of branch cases that had been taken to reach this point in symbolic
execution. A branch path can optionally be passed to the continuation function,
which would then be used to select the next state configuration. Passing no
branch path preserves original behaviour, popping the first available state
config.

This implementation required a change to the \texttt{cont\_func} type for
handling the ends of branches, as the end of a branch requires result data for
use in unifications, but the provided \texttt{Finished} variant gives no
continuation function. A new variant, \texttt{EndOfBranch}, shown in
\autoref{lst:contfunc-type-new} has been introduced to cover the case where
there may still be state configurations left to consider, but none that match
the provided branch path. This variant gives both the results for the relevant
symbolic execution branch, and a cont func that can be used to explore other
branches. The \texttt{Normal} variant has been extended to include information
on the branch path, and any new branch cases resulting from the execution step.

\begin{listing}[!ht]
\noindent\rule{\textwidth}{0.5pt}
\vspace{-0.6cm}
\begin{minted}{ocaml}
type 'a cont_func_f = ?path:branch_path -> unit -> 'a cont_func

and 'a cont_func =
  | Finished of 'a list
  | Continue of
      (Logging.ReportId.t option
      * branch_path
      * branch_case list option
      * 'a cont_func_f)
  | EndOfBranch of 'a * 'a cont_func_f
\end{minted}
\vspace{-0.4cm}
\noindent\rule{\textwidth}{0.5pt}
\vspace{-0.6cm}
\caption{The new \texttt{cont\_func} type, in the \texttt{GInterpreter} module}
\label{lst:contfunc-type-new}
\end{listing}


\subsection{Navigating Symbolic Execution}%
\label{sec:navigating-symex}

Despite a mechanism now being in place for selecting the branch of execution, it
won't be of much use to users without a means of navigating the various paths
through a symbolic execution. To this end, an `execution map', or
\texttt{ExecMap} (\autoref{lst:execmap}) was developed to keep track of that
very process.

\begin{listing}[!ht]
\noindent\rule{\textwidth}{0.5pt}
\vspace{-0.6cm}
\begin{minted}{ocaml}
module ExecMap = struct
  type unifys = (rid * Unifier.unify_kind * UnifyMap.unify_result) list
  [@@deriving yojson]

  type cmd_data = {
    id : rid;
    origin_id : int option; [@name "originId"]
    display : string;
    unifys : unifys;
    errors : string list;
  }
  [@@deriving yojson]

  type 'case t =
    | Nothing
    | Cmd of cmd_data * 'case t
    | BranchCmd of cmd_data * ('case * 'case t) list
    | FinalCmd of cmd_data
  [@@deriving yojson]

  (* ... helper functions ... *)

end
\end{minted}
\vspace{-0.4cm}
\noindent\rule{\textwidth}{0.5pt}
\vspace{-0.6cm}
\caption{The \texttt{ExecMap} module, inside the \texttt{Debugger} module}
\label{lst:execmap}
\end{listing}

After each call of the continuation function, the relevant information from
the executed step is added to the \texttt{ExecMap} in the debugger's internal
state. This map is used at various points in the debugging process, such as
finding the report ID of a command at a particular branch path (and vice versa),
and listing the different branch cases that the interpreter could step to after
a particular command.

The \texttt{ExecMap} can also be supplied wholesale to the debug adapter,
albeit with \texttt{PackagedBranchCase} instead of \texttt{branch\_case}.
The reason for this, and for the \texttt{'case} type variable in
\texttt{ExecMap}, is to abstract away the details of the \texttt{branch\_case}
type for the debugger frontend. \texttt{PackagedBranchCase}
(\autoref{lst:packagedbranchcase}) simply provides the `kind' of branch case,
a pretty-printed form of both the branch case's kind and value, and the JSON
for the \texttt{branch\_case} represented. This allows the frontend to directly
use a representation fit for the user's viewing, while still holding on to the
necessary data structure to talk back to the debugger.

\begin{listing}[!ht]
\noindent\rule{\textwidth}{0.5pt}
\vspace{-0.6cm}
\begin{minted}{ocaml}
module PackagedBranchCase = struct
  type t = {
    kind : string;
    display : string * string;
    json : Yojson.Safe.t;
  }
  [@@deriving yojson]

  (* ... helper functions ... *)

end
\end{minted}
\vspace{-0.4cm}
\noindent\rule{\textwidth}{0.5pt}
\vspace{-0.6cm}
\caption{The \texttt{PackagedBranchCase} type, inside the \texttt{Debugger} module}
\label{lst:packagedbranchcase}
\end{listing}


\subsection{Debugging Unification}%
\label{sec:debug-unify}

As outlined in \autoref{sec:current:dap}, the original debugger implementation
does not perform unification for postconditions at all --- seeing as
postconditions are the only unification guaranteed to be present in any
verification, this is a major blind spot in debugger functionality. To remedy
this, the result information provided in the \texttt{} cont func variant
introduced in \autoref{sec:interpreter-branching} proves its usefulness.
When the debugger encounters the end of a branch, it passes the execution
result to a function in the newly introduced \texttt{Verifier.Debug} module
(\autoref{lst:verifier-debug}). This module was the result of bisecting the
unification process, refactoring the part of the code that actually performs
unification to its own function, allowing the debugger to use only the parts
of the process that it needs to.

\begin{listing}[!ht]
\noindent\rule{\textwidth}{0.5pt}
\vspace{-0.6cm}
\begin{minted}{ocaml}
module Debug : sig
val get_tests_for_prog : prog_t -> proc_tests

val analyse_result :
  t -> Logging.ReportId.t -> SAInterpreter.result_t -> bool
end
\end{minted}
\vspace{-0.4cm}
\noindent\rule{\textwidth}{0.5pt}
\vspace{-0.6cm}
\caption{The signature of the \texttt{Verifier.Debug} module}
\label{lst:verifier-debug}
\end{listing}

Now that unification was actually being performed where necessary, the debugger
(and the frontend, eventually) needs a way of navigating the structure of
unification; the \texttt{UnifyMap} (\autoref{lst:unifymap}) serves this purpose.
This structure is similar to the \texttt{ExecMap}, albeit simpler, as branches
only occur at the start of unification, and the equivalent of a \texttt{Nothing}
variant wasn't necessary since unification happens all at once. When
postcondition unification is manually performed, or another unification (such as
for a function call) is requested, the representative \texttt{UnifyMap} is
constructed and cached in the debugger state.

\begin{listing}[!ht]
\noindent\rule{\textwidth}{0.5pt}
\vspace{-0.6cm}
\begin{minted}{ocaml}
module UnifyMap = struct
  open Verification.SUnifier.Logging

  type unify_kind = Unifier.unify_kind [@@deriving yojson]
  type unify_result = Success | Failure [@@deriving yojson]

  type assertion_data = {
    id : rid;
    fold : (rid * unify_result) option;
    assertion : string;
    substitutions : (string * string) list;
  }
  [@@deriving yojson]

  type unify_seg =
    | Assertion of assertion_data * unify_seg
    | UnifyResult of rid * unify_result
  [@@deriving yojson]

  type map = Direct of unify_seg | Fold of unify_seg list
  [@@deriving yojson]

  type t = unify_kind * map [@@deriving yojson]

  (* ... helper functions ... *)

end
\end{minted}
\vspace{-0.4cm}
\noindent\rule{\textwidth}{0.5pt}
\vspace{-0.6cm}
\caption{The types of the \texttt{Debugger.UnifyMap} module}
\label{lst:unifymap}
\end{listing}

With this in place, the debugger has all the necessary information at its
disposal necessary to allow an interface to deep dive into the unification
process.
